---
title: "Examples of using ggplot"
author: "Rashedul"
date: "7/2/2019"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### load libraries

```{r}
#install.packages("gapminder")
#install.packages("tidyverse")
library(tidyverse)
library(gapminder)
```

### check data

```{r}
head(gapminder)
```

### make scatter plot 

```{r}
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point()

```

### log transformation of values

```{r}
#
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point() +
  scale_x_log10()

# try another way
ggplot(gapminder, aes(log10(gdpPercap), lifeExp)) +
  geom_point()
```

### change color in scatter plot

```{r}
ggplot(gapminder, aes(log10(gdpPercap), lifeExp)) +
  geom_point(aes(color=continent))

#change sige of points
ggplot(gapminder, aes(log10(gdpPercap), lifeExp)) +
  geom_point(aes(color=continent), size = 3)
```

### change some other parameters in scatter plot

```{r}
ggplot(gapminder, aes(log10(gdpPercap), lifeExp)) +
  geom_point(pch=17, size=2, alpha= .8, aes(color=continent)) 

# what about this
ggplot(gapminder, aes(log10(gdpPercap), lifeExp)) +
  geom_point(pch=17, size=2, alpha= .8, color= "red")

```

### use smoothing function

```{r}
ggplot(gapminder, aes(log10(gdpPercap), lifeExp)) +
  geom_point() +
  geom_smooth()
```

### specify the method for smoothing

```{r}
ggplot(gapminder, aes(log10(gdpPercap), lifeExp)) +
  geom_point() +
  geom_smooth(lwd=2, se=FALSE, method="lm", col="blue")

# smooth by continent
ggplot(gapminder, aes(log10(gdpPercap), lifeExp)) +
  geom_point() +
  geom_smooth(aes(color = continent))

# smooth by continent
ggplot(gapminder, aes(log10(gdpPercap), lifeExp)) +
  geom_point(aes(color = continent)) +
  geom_smooth(aes(color = continent))
```

### example on faceting

```{r}
ggplot(gapminder, aes(log10(gdpPercap), lifeExp)) +
  geom_point() +
  facet_wrap(~continent)
```

### change faceting  by column

```{r}
ggplot(gapminder, aes(log10(gdpPercap), lifeExp)) +
  geom_point() +
  facet_wrap(~continent, ncol=1)
```

### change faceting by row

```{r}
ggplot(gapminder, aes(log10(gdpPercap), lifeExp)) +
  geom_point() +
  facet_wrap(~continent, nrow =1)
```

### boxplot and jitter

```{r}
#boxplot
ggplot(gapminder, aes(continent, lifeExp)) +
  geom_jitter() +
  geom_boxplot()

#alpha
ggplot(gapminder, aes(continent, lifeExp)) +
  geom_jitter(alpha=1/2) +
  geom_boxplot()

#fill color
ggplot(gapminder, aes(continent, lifeExp)) +
  geom_jitter(alpha=1/2) +
  geom_boxplot(aes(fill = continent))

```

### violin and jitter

```{r}
ggplot(gapminder, aes(continent, lifeExp)) +
  geom_jitter() +
  geom_violin()

#alpha
ggplot(gapminder, aes(continent, lifeExp), fill = continent) +
  geom_jitter(alpha=1/2) +
  geom_violin()

#color
ggplot(gapminder, aes(continent, lifeExp), fill = continent) +
  geom_jitter(alpha=1/2) +
  geom_violin(aes(fill = continent))
```

### histograms

```{r}
ggplot(gapminder, aes(lifeExp)) +
  geom_histogram()

# change binsize
ggplot(gapminder, aes(lifeExp)) +
  geom_histogram(bins=50)

#
ggplot(gapminder, aes(lifeExp)) +
  geom_histogram(bins=100)

#
ggplot(gapminder, aes(lifeExp)) +
  geom_histogram(bins=500) 

# color by continent
ggplot(gapminder, aes(lifeExp)) + 
  geom_histogram(aes(color=continent))

#fill
ggplot(gapminder, aes(lifeExp)) + 
  geom_histogram(aes(fill=continent))
```

### density plot

```{r}
ggplot(gapminder, aes(lifeExp)) + 
  geom_density()

# color
ggplot(gapminder, aes(lifeExp)) + 
  geom_density(aes(fill= "red"))
```

### change alpha (transparancey)

```{r}
ggplot(gapminder, aes(lifeExp)) + 
  geom_density(aes(fill= "red"), alpha=1/4)

#
ggplot(gapminder, aes(lifeExp)) + 
  geom_density(aes(fill=continent), alpha=1/4)
```

### density and histogram

```{r}
ggplot(gapminder, aes(lifeExp)) + 
    geom_density(size=1.5, fill="pink", alpha=0.5) +
    geom_histogram(aes(y=..density..), binwidth=4, color="black", fill="blue", alpha=0.5)
```

### time series line plot

```{r}
ggplot(gapminder, aes(x=year, y=lifeExp, group=country)) +
    geom_line()
```

### line plot for median life exp

```{r}
gapminder %>%
    group_by(continent, year) %>%
    summarise(lifeExp=median(lifeExp)) %>%
    ggplot(aes(x=year, y=lifeExp, color=continent)) +
     geom_line(size=1) + 
     geom_point(size=1.5) 
```

### line plot for median life exp and smooth

```{r}
gapminder %>%
    group_by(continent, year) %>%
    summarise(lifeExp=median(lifeExp)) %>%
    ggplot(aes(x=year, y=lifeExp, color=continent)) +
     geom_line(size=1) + 
     geom_point(size=1.5) +
    geom_smooth(aes(fill=continent), method="lm")
```





